---
title: rmapzen
output: 
    github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

## Introduction

NOTE: [Mapzen has shut down](https://mapzen.com/blog/shutdown/). However, the rmapzen package can be used with [other providers](https://mapzen.com/blog/migration/). The hosting service can be updated via `options`, for example, to update the search host (for `mz_search`, `mz_geocode`, and `mz_structured_search`):

```{r, eval = FALSE}
options(RMAPZEN_SEARCH_HOST = "api.geocode.earth")
```

-----------------------------------

`rmapzen` is a client for the Mapzen API. For an introduction, detailed examples, and installation instructions, see: [https://tarakc02.github.io/rmapzen/](https://tarakc02.github.io/rmapzen/). The package is available [on CRAN](https://cran.r-project.org/package=rmapzen), to install:

```{r, eval = FALSE}
install.packages("rmapzen")
```

For more information about the Mapzen API, see [https://mapzen.com/documentation/](https://mapzen.com/documentation/). 
